Simple connectome inference from partial correlation statistics in calcium imaging

### 2014-06-18 article.tex
intro to section experiment

The full signal processing of our approach is defined by
\begin{align}
w(h(g(f_k(x_i^t))))
\end{align}
where $f_k$ is one of the low-pass filter designs.

@@

@@

We compare our approach to the Pearson correlation, the generalized transfer
entropy (GTE) and GENIE3.

GTE, the baseline of the challenge, measures for all pairs of
neurons $i$ and $j$ the reduction of uncertainty in predicting future values of
$i$ given the observation of the time series of the neuron $j$
conditionally to the observation of past time series of neuron $i$,
whenever the average fluorescence is over a conditioning level.
Conditioning level was set to reproduce baseline of the challenge.

GENIE3 (\cite{huynhthu2010inferring}) translates the inference of the
network of $p$ neurons into $p$ significant variable tasks. For each task,
the fluorescence level of a neuron is predicted from the fluorescence
of all other neurons using a tree based ensemble such as
random forest (\cite{breiman2001random}). The importances
(\cite{louppe2013understanding}) given by the tree ensemble to predict
the time series of a neuron given the others are used to infer
a connectivity score matrix. The following parameters are used
$10000$ trees, $\log_2{p}$ neurons are randomly selected at each tree
node to split the data and the maximal tree depth is constrained
to 1, 2 or 3.

@@

@@

@@

\begin{table}
\small
\centering
\begin{tabular}{@{}l *{8}{c}@{}}
  & \multicolumn{4}{c}{AUROC} & \multicolumn{4}{c}{AUPRC} \\
\cmidrule(r){2-5}\cmidrule(r){6-9}
Methods $\backslash$ normal- & 1 & 2 & 3 & 4 & 1 & 2 & 3 & 4 \\
\hline \hline
% Directivity \textcolor{red}{appendix?} & 0.535 & 0.526 & 0.533 & 0.535 & 0.012 & 0.012 & 0.012 & 0.012 \\
Correlation            & 0.886 & 0.884 & 0.891 & 0.877 & 0.153 & 0.145 & 0.170 & 0.132 \\
GENIE3 (depth = 1)     & 0.878 & & & & 0.184 & & & \\
GENIE3 (depth = 2)     & 0.889 & & & & 0.217 & & & \\
GENIE3 (depth = 3)     & & & & & & & & \\
GTE                    & 0.890 & 0.893 & 0.894 & 0.873 & 0.171 & 0.174 & 0.197 & 0.142 \\
Partial correlation    & 0.930 &  0.927 &  0.926 &  0.923& 0.363  & 0.347 &  0.346 & 0.328 \\
Partial correlation (opt.) & 0.932 & 0.931 & 0.930 & 0.928 & 0.364 & 0.366 & 0.359 & 0.344 \\
Our approach (opt. \& avg.)    & 0.943 & 0.942 & 0.942 & 0.939 & 0.403 & 0.404 & 0.398 & 0.388 \\
% Our approach + dir  \textcolor{red}{appendix?}   & 0.944 & 0.943 & 0.942 & 0.940 & 0.404 & 0.405 & 0.399 & 0.389 \\
\end{tabular}
\caption{TODO}
\label{tab:tab1}
\end{table}

>>>>

The full signal processing of our approach is defined by the composed function
$
w \cdot h \cdot g \cdot f_k 
$
where $f_k$ is one of the low-pass filter designs.

@@

\textcolor{red}{SEUIL 0.11 for tabulars!!!}

In this section, we will first describe datasets and methods used for the comparison. Then, on Tab.\ref{tab:tab3}, our approach is decomposed to show the contribution of each stages . Afterwards, we will perform a comparison between several methods on Tab.\ref{tab:tab1} (or Fig.?). Note that two metrics are used to evaluate the performance. The end of this section is devoted to a discussion about the choice of metric and its relevance for the challenge.

@@

@@

\textcolor{red}{à vérifier} GTE is a method which characterizes a relationship
between neurons by an estimation of the generalized transfer entropy
\citep{stetter2012model}. The goal of this method tries to see if the next
time step value is better from the previous values of this time series or from
time series of other neurons. In Tab.\ref{tab:tab1}, default parameters were
used to infer networks.

GENIE3 is a tree-based method which sees the network inference problem as $p$,
the number of variables, independent supervised learning problems (todo cite:
Irrthum?Van Anh?Pierre?). Formally, each sub-problem consists in the
prediction of one variable, the target, from the remaining $p-1$ variables.
This method uses random forests to estimate the target variable and variable
importances can be derived to estimate the adjacency matrix $A$ (see (todo
cite) for a more detailed description of the inference method). We compute
$10000$ trees for each subproblem with the randomization parameter equals to
$\log_2{p}$ and we limite the tree max-depth to respectively $1$, $2$ and $3$.

@@

\textcolor{red}{TODO : Discussion depth}

@@

\begin{table}[tbh]
\centering
\begin{tabular}{@{}l *{8}{c}@{}}
\hline
  & \multicolumn{4}{c}{AUROC} & \multicolumn{4}{c}{AUPRC} \\
\hline
Methods $\backslash$ normal- & 1 & 2 & 3 & 4 & 1 & 2 & 3 & 4 \\
No  filtering       & 0.777 & 0.767 & 0.772 & 0.774 & 0.070 & 0.064 & 0.068 & 0.072\\
P                   & 0.925 & 0.925 & 0.924 & 0.923 & 0.322 & 0.324 & 0.323 & 0.312\\
P + W               & 0.931 & 0.930 & 0.928 & 0.927 & 0.333 & 0.334 & 0.327 & 0.313\\
P + W + PCA         & 0.932 & 0.931 & 0.930 & 0.928 & 0.364 & 0.366 & 0.359 & 0.344\\
Averaging           & 0.943 & 0.942 & 0.942 & 0.939 & 0.403 & 0.404 & 0.398 & 0.388\\
Averaging + Directivity & 0.944 & 0.943 & 0.942 & 0.940 & 0.404 & 0.405 & 0.399 & 0.389\\
\end{tabular}
\caption{TODO \textcolor{red}{version simple!!}}
\label{tab:tab3}
\end{table}

\begin{table}
\small
\centering
\begin{tabular}{@{}l *{8}{c}@{}}
  & \multicolumn{4}{c}{AUROC} & \multicolumn{4}{c}{AUPRC} \\
\hline \hline
Methods $\backslash$ normal- & 1 & 2 & 3 & 4 & 1 & 2 & 3 & 4 \\
\hline \hline
% Directivity \textcolor{red}{appendix?} & 0.535 & 0.526 & 0.533 & 0.535 & 0.012 & 0.012 & 0.012 & 0.012 \\
Correlation            & 0.886 & 0.884 & 0.891 & 0.877 & 0.153 & 0.145 & 0.170 & 0.132 \\
GENIE3 (depth = 1)     & 0.878 & & & & 0.184 & & & \\
GENIE3 (depth = 2)     & 0.889 & & & & 0.217 & & & \\
GENIE3 (depth = 3)     & & & & & & & & \\
GTE                    & 0.890 & 0.893 & 0.894 & 0.873 & 0.171 & 0.174 & 0.197 & 0.142 \\
Partial correlation    & 0.930 &  0.927 &  0.926 &  0.923& 0.363  & 0.347 &  0.346 & 0.328 \\
Partial correlation (avg.) & 0.938 & & & 0.932& 0.391 & & & 0.374\\
% Our approach simple + dir & 0.939 & & & 0.933& 0.392 & & & 0.375\\
Partial correlation (opt.) & 0.932 & 0.931 & 0.930 & 0.928 & 0.364 & 0.366 & 0.359 & 0.344 \\
Our approach (opt. \& avg.)    & 0.943 & 0.942 & 0.942 & 0.939 & 0.403 & 0.404 & 0.398 & 0.388 \\
% Our approach + dir  \textcolor{red}{appendix?}   & 0.944 & 0.943 & 0.942 & 0.940 & 0.404 & 0.405 & 0.399 & 0.389 \\
\end{tabular}
\caption{TODO}
\label{tab:tab1}
\end{table}
